Mouse Data Overview 1
========================================================
```{r options}
opts_chunk$set(echo=FALSE, message=FALSE)
```

```{r setup, results='hide'}
library(cummeRbund)
library(xtable)
library(limma)
library(GSA)
library(gplots)
library(marray)
library(ggplot2)

analysisdir<-"/n/rinn_data1/users/agroff/GITHUB/BrainMap/analysis/"
diffdir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs"
GTF<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/annotation/mm10_gencode_vM2_with_lncRNAs_and_LacZ.gtf"

dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Celrr_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks(gtfFile=GTF,genome="mm10")

alpha<-0.05



setwd(diffdir)
names<-list.files()
names<-gsub("-","",names)
setwd(analysisdir)
diffGeneSummary<-data.frame(t(names), row.names=NULL)
colnames(diffGeneSummary)<-names
diffGeneSummary$unit_test<-NULL
diffGeneSummary$init<-NULL

```

# Adult Samples 
- All have strange density plots: add pseudocount 

Without Pseudocount

```{r density}
csDensity(genes(cuff),replicates=TRUE)
```

With Pseudocount = 0.01 

```{r density pseudo}
csDensity(genes(cuff),pseudocount=0.01,replicates=TRUE)
```


## ADULT WILDTYPES:

### Genotyping
- wt 12 is ko/het (JR817)

```{r genotyping}
genotypingGeneIDs<-c("Celrr","LacZ","Eif2s3y")
genotypingGenes<-getGenes(cuff,genotypingGeneIDs)
csHeatmap(genotypingGenes,replicates=T)
```

### QC 
- wts 9, 10 have smaller fpkm distribution (replicates)

```{r boxplot replicates}
csBoxplot(genes(cuff),replicates=T)
```

- isoform fpkm distributions: WT smaller on whole, but replicates look fine? 

Boxplot (isoforms)

```{r boxplot isoforms}
csBoxplot(isoforms(cuff))
```

Boxplot (isoforms, replicates)

```{r boxplot isoforms replicates}
csBoxplot(isoforms(cuff), replicates=T)
```


### Other Notes

## ADULT KNOCKOUT STRAINS 

### Celrr
```{r adult celr diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Celrr_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Celrr_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Crnde

- Either wrong genotyping or isoform beyond KO region stabilized by LZ (need to see tracks)

```{r crnde adult}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Crnde_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()
id<-"Crnde"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```


- Wrong genotyping CRNDE_2 (JR818)

```{r crnde lz adult}
id<-"LacZ"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c("Crnde","LacZ")
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```

```{r adult crnde diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Crnde_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`

### Eldr

- Either wrong genotyping or isoform beyond KO region stabilized by LZ (need to see tracks)

```{r eldr adult}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Eldr_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()
id<-"Eldr"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)

```


```{r eldr lz adult}
id<-"LacZ"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c("Eldr","LacZ")
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```

Problem child is Eldr_0 (JR733)
Does Eldr_1 look like a het? (JR799)

```{r adult Eldr diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Eldr_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`

### Haunt

```{r adult haunt diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Haunt_vs_WT_Adult/"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Haunt_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Kantr (Jarid1C)
YAY! They actually cluster!
```{r Kantr adult}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Kantr_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()

csDendro(genes(cuff),replicates=TRUE)

id<-"Kantr"
myGene<-getGene(cuff,id)

```

```{r adult Kantr diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Kantr_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`

### Brn1a
3 isoforms expressed

```{r adult Brn1a diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Brn1a_vs_WT_Adult/"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$lincBrn1a_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Brn1b
LacZ expressed slightly higher than lincRNA

```{r adult lincBrn1b diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Brn1b_vs_WT_Adult/"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$lincBrn1b_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### linc-Cox2
YAY! They cluster together! (is this a litter thing? sex difference thing?) -- Weird, in the other one they cluster together. Diff w/ having mm10/gtf specified? 


```{r lincCox2 adult}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Cox2_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()

csDendro(genes(cuff),replicates=TRUE)

id<-"linc-Cox2"
myGene<-getGene(cuff,id)

```

**THIS DENDROGRAM LOOKS DIFFERENT WHEN PRINTED IN R STUDIO!**

Endogenous linc and LZ approx same expression levels 

```{r adult lincCox2 diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Cox2_vs_WT_Adult/"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$lincCox2_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### linc-Enc1a

I think this one MIGHT be a mapping error. Endogenous linc expressed at ~ WT levels (very low), but all cells have lacZ robustly expressed. Potentially hets I guess 

```{r lincEnc1a}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Enc1_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()
id<-"linc-Enc1"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r lincEnc1a lz adult}
id<-"LacZ"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c("linc-Enc1","LacZ")
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```

```{r adult lincEnc1a diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$lincEnc1_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Manr
Same as with linc-Enc1a, may be extra-knockout region OR perhaps hets or some sort of artifact.. 

```{r MANR}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Manr_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()
id<-"Manr"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r MANR lz adult}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```

```{r adult Manr diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Manr_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Peril

Fairly certai this is extra-knockout region. Need to make tracks! 

```{r Peril}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Peril_vs_WT_Adult"
setwd(dir)
cuff<-readCufflinks()
id<-"Peril"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r Peril lz adult}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```

```{r adult Peril diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Peril_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`



### Trp53cor1 (linc-p21)
Everything very lowly expressed 

```{r adult Trp53cor diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Trp53cor1_vs_WT_Adult/"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Trp53cor1_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Tug1

LacZ expressed far lower than Tug1 lincRNA 
Why is one of the tug1 replicates missing from plots/dendrogram??

```{r adult Tug1 diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Tug1_vs_WT_Adult/"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Tug1_vs_WT_Adult<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


# Embryonic Samples

- wt13 

## WTS:

### Genotyping

### QC 

### Other Notes

## KO Strains

### Celrr
- Celrr: celrr 0 is not KO (seems WT) JR769

```{r celrr embryonic}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Celrr_vs_WT_Embryonic"
setwd(dir)
cuff<-readCufflinks()
id<-"Celrr"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r Celrr lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```


```{r Celrr embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Celrr_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`

Cell compartment: lots of cytoplasmic stuff... is this linc cytoplasmic?? 

### Crnde
- actually cluster together!

```{r Crnde embryonic dendro}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Crnde_vs_WT_Embryonic"
setwd(dir)
cuff<-readCufflinks()
csDendro(genes(cuff),replicates=TRUE)
```


Endogenous expression (isoforms)

```{r Crnde embryonic}
id<-"Crnde"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap
Crnde_2 may be het? Need to check tracks! 

```{r Crnde lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```

```{r Crnde embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Crnde_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`

-ECM/collagen... 

### Eldr

genotyping looks good!

Kegg: huntingtons/alzheimers/cancer

```{r Eldr embryonic dendro}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Eldr_vs_WT_Embryonic"
setwd(dir)
cuff<-readCufflinks()
```

```{r Eldr embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Eldr_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Haunt
digital genotyping looks good 

```{r Haunt embryonic dendro}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Haunt_vs_WT_Embryonic"
setwd(dir)
cuff<-readCufflinks()
```

```{r Haunt embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Haunt_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Kantr
Yay! Cluster together! 

```{r Kantr embryonic dendro}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Kantr_vs_WT_Embryonic"
setwd(dir)
cuff<-readCufflinks()
csDendro(genes(cuff),replicates=TRUE)

id<-"Kantr"
myGene<-getGene(cuff,id)
```

Endogenous Kantr expression: 

```{r Kantr embryonic}

myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r Kantr lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```

```{r Kantr embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Kantr_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`

### Brn1a -- NOT DONE YET 
```{r brn1a}
diffGeneSummary$lincBrn1a_vs_WT_Embryonic<-NULL
```

### Brn1b 
- digital genotyping looks good 

```{r Brn1b embryonic diff}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Brn1b_vs_WT_Embryonic"
setwd(dir)
cuff<-readCufflinks()
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$lincBrn1b_vs_WT_Embryonic<-length(sig)
```
Number of differentially expressed genes:`r length(sig)`


```{r GO setup}

library(ReactomePA)
library(DOSE)

sigGeneIDs<-getSig(cuff, alpha=alpha)
sigGenes<-getGenes(cuff,sigGeneIDs)
geneAnnot<-annotation(sigGenes)
geneNames<-geneAnnot$gene_short_name
sigDiff<-diffData(sigGenes)

```


```{r GO analysis setup and functions}
library(clusterProfiler)
require(biomaRt)

ensembl <- useMart("ENSEMBL_MART_ENSEMBL","mmusculus_gene_ensembl", 
host="www.ensembl.org")

#Get entrezIDs
getEntrezIDs<-function (geneNames)
{
    tmp <- getBM(attributes = c("entrezgene"), filters = "mgi_symbol", values = geneNames, mart = ensembl)
    tmp
}
sigEntrez<-getEntrezIDs(geneNames)
sigEZ<-strsplit(as.character(sigEntrez), ", ")
sigEZ<-unlist(sigEZ)
require(org.Mm.eg.db) #mapping of entrez IDs to genbank

goBP<-enrichGO(gene=sigEZ, organism="mouse",ont="BP",pvalueCutoff=0.01,readable=T)

goMF<-enrichGO(gene=sigEZ, organism="mouse",ont="MF",pvalueCutoff=0.01,readable=T)

goCC<-enrichGO(gene=sigEZ, organism="mouse",ont="CC",pvalueCutoff=0.01,readable=T)

kegg<-enrichKEGG(gene=sigEZ, organism="mouse",pvalueCutoff=1, readable=T)

pathway<-enrichPathway(gene=sigEZ,organism="mouse",pvalueCutoff=0.01, readable=T)

```

```{r GO figures,  fig.height=5, fig.width=8}
plot(goBP,showCategory=10) + theme(axis.text.x=element_text(angle=-90,hjust=0)) + ggtitle("Enriched BP")

plot(goMF,showCategory=10) + theme(axis.text.x=element_text(angle=-90,hjust=0)) + ggtitle("Enriched MF")

plot(goCC,showCategory=10) + theme(axis.text.x=element_text(angle=-90,hjust=0)) + ggtitle("Enriched CC")

plot(kegg, showCategory=10) + ggtitle("Kegg Pathways")

plot(pathway,showCategory=10) + ggtitle("Reactome pathway enrichment")
detach("package:biomaRt")
```

### linc-Cox2

Digital genotyping looks good! 

GO terms ribosome/translation related... 

Cluster together! Yay! 

```{r lincCox2 embryonic dendro}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Cox2_vs_WT_Embryonic/"
setwd(dir)
cuff<-readCufflinks()
csDendro(genes(cuff),replicates=TRUE)

id<-"linc-Cox2"
myGene<-getGene(cuff,id)
```


```{r lincCox2 embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$lincCox2_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### linc-Enc1a

```{r lincEnc1a embryonic}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/linc-Enc1_vs_WT_Embryonic/"
setwd(dir)
cuff<-readCufflinks()
id<-"linc-Enc1"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r lincEnc1a lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```


```{r lincEnc1a embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$lincEnc1_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Manr

- GO terms lots of ion/lipid/protein binding; kegg neuroactive ligand-receptor interaction; and reactome has interesting categories but graph is messed up. 

- Hets? Need to see track... 

```{r Manr embryonic}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Manr_vs_WT_Embryonic/"
setwd(dir)
cuff<-readCufflinks()
id<-"Manr"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r Manr lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```


```{r Manr embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Manr_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Peril

Pretty sure this is an exon-stabilized-by-lacZ issue, but need to see tracks!

```{r Peril embryonic}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Peril_vs_WT_Embryonic/"
setwd(dir)
cuff<-readCufflinks()
id<-"Peril"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r Peril lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```


```{r Peril embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Peril_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


### Trp53cor1 (linc-p21)

Trp53cor1_1 is wildtype (JR813)

```{r Trp53 embryonic}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Trp53cor1_vs_WT_Embryonic/"
setwd(dir)
cuff<-readCufflinks()
id<-"Trp53cor1"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r Trp53 lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```


```{r Trp53 embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Trp53cor1_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`

### Tug1
Digital genotyping looks good. 
Tug1 far more highly expressed than lacZ. 

```{r tug1 embryonic}
dir<-"/n/rinn_data1/seq/lgoff/Projects/BrainMap/data/diffs/Tug1_vs_WT_Embryonic/"
setwd(dir)
cuff<-readCufflinks()
id<-"Tug1"
myGene<-getGene(cuff,id)
expressionBarplot(isoforms(myGene), replicates=T)
```

LacZ and genotyping heatmap

```{r tug1 lz embryonic}
lz<-"LacZ"
myGene<-getGene(cuff,lz)
#expressionBarplot(isoforms(myGene), replicates=T)
ids<-c(id,lz)
myGenes<-getGenes(cuff,ids)
csHeatmap(myGenes,replicates=TRUE)
```


```{r tug1 embryonic diff}
sig<-getSig(cuff, alpha=alpha)
diffGeneSummary$Tug1_vs_WT_Embryonic<-length(sig)
```

Number of differentially expressed genes:`r length(sig)`


# SUMMARY

**list of samples to re-check or exclude**

GENOTYPING:
- Adult:
  - WT 12 (JR817)
  - CRNDE_2 (JR818)
  - Eldr_0 (JR733) 
  - Does Eldr_1 look like a het? (JR799)
- Embryonic:
  - WT 13 
  - celrr 0 is not KO (seems WT) JR769
  - check crnde? (esp #2)
  - check Manr(#2?)
  - linc-P21_1 is wt not ko 

enc1a problem child -- need to see tracks

QC: adult wt 9, 10

Range of number of differentially expressed genes:

```{r summary, results='asis'}
print(xtable(summary(t(diffGeneSummary))), type="html")
```


Number of differentially expressed genes, by condition: 

```{r diff summary}
x<-as.numeric(diffGeneSummary[1,])
xx<-rbind(x,x)
colnames(xx)<-gsub("_vs_WT","",colnames(diffGeneSummary))
colnames(xx)<-gsub("Embryonic","E",colnames(xx))
colnames(xx)<-gsub("Adult","A",colnames(xx))
xx_ordered<-xx[,order(colnames(xx))]
rownames(xx_ordered)<-c("","")
heatmap.2(xx_ordered, Rowv=FALSE, breaks=15, dendrogram="none", Colv=FALSE, sepwidth=c(.9,.9),trace="none", col=terrain.colors, lhei=c(.5,.5))

barplot(x)
#axes=TRUE,names.arg=colnames(diffGeneSummary)
```

**Interesting or promising strains!**
- Kantr!


# TO DO
- tracks!
- overlap!
- finish tidying script + reprint!
- fix gsea printing 
