`r strain` KO vs WT (Embryonic)
======================================

# questions
- overdispersion
- get tables to work
- counts


```{r init, echo=FALSE}
# Set knitr opts
path<-paste("figure/",strain,"/", sep="")
opts_chunk$set(fig.path=path)


### THINGS TO CACHE GO HERE AND NOWHERE ELSE 




```

```{r vars_and_setup,echo=FALSE,results='hide',message=FALSE}
library(cummeRbund)
library(xtable)
library(limma)
library(GSA)
library(gplots)
library(marray)

reactome_gs <- GSA.read.gmt("/n/rinn_data1/users/agroff/GSEA/c2.cp.reactome.v4.0.symbols.gmt")
biocarta_gs <- GSA.read.gmt("/n/rinn_data1/users/agroff/GSEA/c2.cp.biocarta.v4.0.symbols.gmt")

# Constant variables
alpha<-0.05

```

# Intialize
```{r cummeRbund and data setup, echo=FALSE}
setwd(dir)
cuff<-readCufflinks()
reps<-replicates(cuff)
files<-lapply(reps$file,function(x){strsplit(x, "/")})

#field 9 will be JR number 
files<-as.data.frame(files)
samples<-as.data.frame(t(files[9,]))

#look up strain, litter, sex, notes in master sheet 

LINCRNA<-strain

```

# Design Overview

This file shows the wt-v-ko comparison for `r strain`. 

Samples used are:
```{r}
samples
```


Cuff overview:
```{r}
cuff
```

## Replicates (MAKE TABLES PRETTY >:| )
```{r replicates, results='asis'}
print(xtable(replicates(cuff)),type="html")
```

# QC

## Dispersion

This shows count (WHAT IS COUNT? fragment counts? per gene?) vs dispersion (of those counts), or the spread in measurements for a particular gene across replicates. 

from cummerbund manual: Overdispersion is a common problem in RNA-Seq data. As of cufflinks v2.0 mean counts, variance, and dispersion are all emitted, allowing you to visualize the estimated overdispersion for each sample as a quality control measure. - See more at: http://compbio.mit.edu/cummeRbund/manual_2_0.html#sthash.9YRYxwRV.dpuf


```{r dispersion}
dispersionPlot(genes(cuff))

csDendro(genes(cuff),replicates=TRUE)


```
